\textbf{ Interpretable Gaussian Processes for Stellar Variability }

### 2020-12-04 tex/gitlinks.tex
more text

\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/b705a027358ae5fc187985032ec1bcb39175deee/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/b705a027358ae5fc187985032ec1bcb39175deee/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/b705a027358ae5fc187985032ec1bcb39175deee/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

>>>>

\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/c50415978a292c49f67eb6752e3f57c606f8bf6f/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/c50415978a292c49f67eb6752e3f57c606f8bf6f/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/c50415978a292c49f67eb6752e3f57c606f8bf6f/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-04 tex/ms.tex
more text

% Title
\title{%
    \textbf{
        Interpretable Gaussian Processes for Stellar Light Curves
    }
}

@@

\section{Introduction}
\label{sec:intro}
\xxx{Talk about starry, gps, etc.}

\section{Motivation}
\label{sec:motivation}

@@

>>>>

% Title
\title{%
    \textbf{
        Interpretable Gaussian Processes for Stellar Variability
    }
}

@@

\section{Introduction}
\label{sec:intro}

Over the past decade, Gaussian processes (GPs) have gained traction as
a leading tool for modeling correlated signals in astrophysical timeseries.
Despite whatever mystique the words ``Gaussian process'' may evoke, a GP
is nothing but a Gaussian distribution in many (formally infinite)
dimensions. Specifically, it is a Gaussian distribution over
\emph{functions} spanning a continuous domain (in our case, the time domain).
Like a multivariate
Gaussian, a GP is fully specified by a $(K \times 1)$ vector $\pmb{\mu}$ characterizing
the mean of the process and a $(K \times K)$
covariance matrix $\pmb{\Sigma}$ characterizing how the data points
are correlated with each other. To say that a measurement $\mathbf{f}$ is
``distributed as a GP'' means that we may write
%
\begin{align}
    \mathbf{f} \sim \mathcal{N}\left( \pmb{\mu}, \pmb{\Sigma} \right)
    \quad,
\end{align}
%
i.e., the measurement is modeled as a random draw from a multivariate Gaussian.
GPs are therefore extremely easy to sample from.%
\footnote{Given a 1-d array \texttt{mean} and a 2-d array \texttt{cov} in \Python,
    sampling from the corresponding GP (if it exists)
    can be done in a single line of code by calling
    \texttt{numpy.random.multivariate\_normal(mean, cov)}.}
But the real showstopper is the application of GPs to inference problems.
Multivariate Gaussian distributions have a closed-form (marginal) likelihood
function, so it is easy (and extremely efficient) to compute the probability
of one's data conditioned on a given value of $\pmb{\mu}$ and $\pmb{\Sigma}$
(i.e., the ``likelihood''). This can in turn be maximized
to infer the optimal values of the model parameters
or used in a
numerical sampling scheme to compute the probability of those parameters
given the data (i.e., the ``posterior'').
Thanks to modern computer architectures, linear algebra packages, and
GP algorithms,
evaluating the GP likelihood may typically be done in a fraction of a second
for a reasonably-sized dataset (i.e., $K \lesssim 10^5$ datapoints).

Another big advantage of GPs is their flexibility. GPs are often dubbed
a class of ``non-parametric'' models, given that nowhere in the specification
of the GP is there an explicit functional form for $\mathbf{f}$. Rather, GPs are a stochastic
process whose draws can in principle take on \emph{any} functional form,
subject, however, to certain smoothness and correlation criteria
of tunable strictness
that are fully
encoded in the covariance $\pmb{\Sigma}$. A common approach to constructing
$\pmb{\Sigma}$ is to define a continuous \emph{kernel function} $k(t_i, t_j)$
such that the covariance between points at times $t = t_i$ and $t = t_j$ is
%
\begin{align}
    \Sigma_{i,j} = k(t_i, t_j)
    \quad.
\end{align}
%
In many applications, it is customary to further restrict the problem by
assuming that the process is \emph{stationary}, such that we may write
%
\begin{align}
    \Sigma_{i,j} & = k(t_i, t_j)
    \nonumber                                    \\
                 & = k(\left| t_i - t_j \right|)
    \nonumber                                    \\
                 & \equiv k(\Delta t)
    \quad.
\end{align}
%
A stationary process is one that is independent of phase (or, in this case,
the actual value of the time $t$); rather, it depends only on the \emph{difference}
between the phases of two data points. The kernel of a stationary process is
therefore a one-dimensional function, which is typically chosen from a set of
standard functions with desirable smoothness and spectral properties. These
include, for example, the squared exponential kernel
%
\begin{align}
    k_\mathrm{E^2}(\Delta t) = \sigma^2 e^{-\frac{\Delta t}{2\tau}}
    \quad,
\end{align}
%
the Mat\'ern-3/2 kernel
%
\begin{align}
    k_\mathrm{M}(\Delta t) = \sigma^2 \left( 1 + \frac{\sqrt{3}\Delta t}{\tau} \right) e^{-\frac{\sqrt{3}\Delta t}{\tau}}
    \quad,
\end{align}
%
the exponential sine squared kernel,
%
\begin{align}
    k_\mathrm{S^2}(\Delta t) = \sigma^2 e^{-\Gamma\sin^2\left( \frac{\pi \Delta t}{P} \right)}
    \quad,
\end{align}
%
the cosine kernel,
%
\begin{align}
    k_\mathrm{C}(\Delta t) = \cos\left( \frac{2\pi \Delta t}{P} \right)
    \quad,
\end{align}
%
or any linear combination or product of those (as well as many other kernels not
listed here). In the expressions above, $\sigma$, $\tau$, $\Gamma$, and $P$ are
the \emph{hyperparameters} of the kernels, scalar quantities that (together with
the assumed functional form for the kernel) fully specify the covariance of the
Gaussian process.

\xxx{More on this soon!}

@@

\subsection{Ensemble Analyses}
\label{sec:ensemble}
